---
title: "Cross Validation"
author: "Courtney Diamond"
date: "2023-11-27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load key packages

```{r}
library(tidyverse)
library(modelr)

set.seed(1)
```

## Nonlinear data and CV

```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - 0.3) ^ 2 + rnorm(100, 0, 0.3)
  )

nonlin_df |> 
  ggplot(aes(x = x, y = y)) +
  geom_point()
```


Do the train/test split. (First by Hand)

```{r}
train_df = sample_n(nonlin_df, 80)

test_df = anti_join(nonlin_df, train_df, by = "id")

train_df |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_point(data = test_df, color = "red")
```

```{r}
linear_mod = lm(y ~ x, data = train_df)
```

Quick visualization of the linear model

```{r}
train_df |> 
  modelr::add_predictions(linear_mod) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = pred))
```

It's not great! Let's try to do a better one. 

```{r}
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
```

```{r}
train_df |> 
  modelr::add_predictions(smooth_mod) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = pred))
```

Much better.  But let's mess around a little bit.

```{r}
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
train_df |> 
  modelr::add_predictions(wiggly_mod) |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(aes(y = pred))
```

IT WIGGLY

AIC and BIC and F-test don't work here- model isn't nested, isn't clear how many degrees of freedom there are. So instead, we'll use the RMSE method.

```{r}
rmse(linear_mod, train_df)
rmse(smooth_mod, train_df)
rmse(wiggly_mod, train_df)
```

So far looks like the linear one is best, but we don't know for sure! Need to check it out on the testing data for out-of-sample prediction accuracy.

```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```

Good thing we tested- our best model switched from being wiggly to smooth. 

